### 2023-11-28 paper/FOUNDRY-TOSEM_JOURNAL/abbreviations.tex
Upload the journal latex project
This is the version submitted. It is tagged as v3.0. The historic of changes can be accessed from the overleaf version.

% English
\newcommand{\aka}{\hbox{\emph{aka}}\xspace}
\newcommand{\cf}{\hbox{\emph{cf.}}\xspace}
\newcommand{\deletia}{\ldots [deletia] \ldots}
\newcommand{\etal}{\hbox{\emph{et al.}}\xspace}
\newcommand{\eg}{\hbox{\emph{e.g.}}\xspace}
\newcommand{\etc}{\hbox{\emph{etc.}}\xspace}
\newcommand{\ie}{\hbox{\emph{i.e.}}\xspace}
\newcommand{\scil}{\hbox{\emph{sc.}}\xspace} %scilicet: it is permitted to know
%\newcommand{\st}{\hbox{\emph{s.t.}}\xspace}
\newcommand{\viz}{\hbox{\emph{viz.}}\xspace} %videlicet: it is permitted to see
\newcommand{\wrt}{\hbox{\emph{w.r.t.}}\xspace}