### 2023-11-28 paper/FOUNDRY-TOSEM_JOURNAL/main.tex

%\documentclass[manuscript,screen]{acmart}
\documentclass[acmtacm,screen]{acmart}
%\documentclass[acmsmall, screen, review=false]{acmart}
%\documentclass[anonymous,print]{acmart}
%\usepackage[true]{anonymous-acm}
\usepackage[nameinlink]{cleveref}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{framed}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{adjustbox}
\input{abbreviations.tex}

% hyperref commands to correct capitalise \autoref labels
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\def\subsubsectionautorefname{Section}
\def\algorithmautorefname{Algorithm}
%github
\usepackage[switch]{lineno}%line number
%\linenumbers
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%\setcopyright{acmcopyright}
%\copyrightyear{2022}
%\acmYear{2022}
%\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a JOURNAL article.
%\acmJournal{JACM}
%\acmVolume{37}
%\acmNumber{4}
%\acmArticle{111}
%\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Comments %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ifthen}
\usepackage{xcolor}

\newboolean{showcomments}
\setboolean{showcomments}{true} % comment this line to deactivate comments
\ifthenelse{\boolean{showcomments}}{
  \newcommand{\nbc}[3]{
    {\colorbox{#3}{\bfseries\sffamily\scriptsize\textcolor{white}{#1}}}%
    {\textcolor{#3}{\sf\small$\blacktriangleright$\textit{#2}$\blacktriangleleft$}}}
  \newcommand{\todo}[1]{\nbc{TODO}{#1}{blue}\xspace}
}{
  \newcommand{\nbc}[3]{}
  \newcommand{\todo}[1]{}
}
\newcommand{\eb}[1]{\nbc{Earl}{#1}{orange}}
\newcommand{\jp}[1]{\nbc{Justyna}{#1}{blue}}
\newcommand{\ls}[1]{\nbc{Leandro}{#1}{red}}
\newcommand\re[1]{\nbc{To Review:}{#1}{green}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% End Comments %%%%%%%%%%%%%%%%%%%%%%%%%

%%% Project Macros
\newcommand{\FOUNDRY}{\textsc{Foundry}\xspace}
\newcommand{\prodscalpel}{\textsc{prodScalpel}\xspace}
\newcommand{\projURL}{\url{utopia.com}}

\begin{document}

\title{Software Product Line Engineering via Software Transplantation }

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Leandro O. Souza}
\email{leandro.souza@ifba.edu.br}
\orcid{0009-0001-6264-2573}
%\orcid{1234-5678-9012}
%\authornotemark[1]
\affiliation{%
  \institution{Federal Institute of Bahia}
 % \streetaddress{Rua Rogério Nunes Dourado, 27}
  \city{Irecê}
  \state{Bahia}
  \country{Brazil}
 % \postcode{44900-000}
}

\author{Earl T. Barr}
\email{E.Barr@ucl.ac.uk}
\orcid{0000-0003-0771-7891}
\affiliation{
    \institution{University College London}
    \city{London}
    \country{UK}
    }

\author{Justyna Petke}
\email{j.petke@ucl.ac.uk}
\orcid{0000-0002-7833-6044}
\affiliation{
    \institution{University College London}
    \city{London}
    \country{UK}
    }

\author{Eduardo S. Almeida}
\email{esa@rise.com.br}
\affiliation{
    \institution{Federal University of Bahia}
    \city{Salvador}
    \state{Bahia}
    \country{Brazil}
    }

\author{Paulo Anselmo M. S. Neto}
\email{paulo.motant@ufrpe.br}
\orcid{0000-0003-0197-8249}
\affiliation{
    \institution{Federal Rural University of Pernambuco}
    \city{Recife}
    \state{Pernambuco}
    \country{Brazil}
    }

%\author{ADD AUTHOR NAMES}
%\email{leandro.souza@ifba.edu.br}
%\affiliation{%
%  \institution{University College London}
  %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
%  \city{London}
%  \country{UK}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

\begin{abstract}
For companies producing related products, a Software Product Line (SPL) is a software reuse method that improves time-to-market and software quality, achieving substantial cost reductions. 
These benefits do not come for free.
It often takes years to re-architect and re-engineer a codebase to support SPL and, once adopted, it must be maintained. Current SPL practice relies on a collection of tools, tailored for different reengineering phases, whose output developers must coordinate and integrate. 
We present \FOUNDRY, a general automated approach for leveraging software transplantation to speed conversion to and maintenance of SPL. 
\FOUNDRY facilitates feature extraction and migration. 
It can efficiently, repeatedly, transplant a sequence of features, implemented in multiple files. 
We used \FOUNDRY to create two valid product lines that integrate features from three real-world systems in an automated way. 
Moreover, we conducted an experiment comparing \FOUNDRY's feature migration with manual effort. We show that \FOUNDRY automatically migrated features across codebases 4.8 times faster, on average, than the average time a group of SPL experts took to accomplish the task.
\end{abstract}

%We show that \FOUNDRY takes, on average, 18\% of the time to reuse features across codebases than a group of SPL experts.
%\end{abstract}
%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
% ADD CCS CONCEPTS

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Software Product Lines, Software Transplantation, Genetic Improvement}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle
%sections
\input{sections/sec01-introduction}
\label{intro}

\input{sections/sec02-motivating_example.tex}
\label{sec:motivating_example}

\input{sections/sec03-approach.tex}
\label{sec:approach}

\input{sections/sec04-implementation.tex}
\label{sec:implementation}

\input{sections/sec06-case_studies.tex}
\label{sec:case_studies}

\input{sections/sec07-experiment.tex}
\label{sec:experiment}

\input{sections/sec05-automated_spl_reengineering.tex}
\label{sec:automated_spl_reengineering}

\input{sections/sec08-related_work.tex}
\label{sec:related_work}

\input{sections/sec09-conclusion.tex}
\label{sec:conclusion} 
%-------------

%\section{Acknowledgments}

%This work was support by the UKRI EPSRC grant no. EP/P023991/1.

\section{Copyright}

For the purpose of open access, the author has applied a Creative Commons Attribution (CC BY) licence to any Author Accepted Manuscript version arising.

%\section{Appendices}

%If your work needs an appendix, add it before the ``\verb|\end{document}|'' command at the conclusion of your source document.

%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio.bib}

%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}
\endinput

%% End of file `main.tex'.